---
title: "NFL Attendance in 2020"
author: "Cason Wight"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

## Prep Work for Analysis

### Load all tools and functions
```{r utils}
source("utils/eda.R")
source("utils/get_data.R")
source("utils/mcmc_diagnostics.R")
source("utils/stan.R")
source("models/bayesian.R")
source("results/calculations.R")
source("results/plots.R")
```

### Load data
```{r data}
attendance <- get_attendance_data() %>% 
  mutate(home_team = str_replace(home_team, "Redskins", "Commanders"))
print(attendance %>% head())

attendance_2020 <- get_2020_attendance_data() %>% 
  mutate(team_name = str_replace(team_name, "Redskins", "Commanders"))
print(attendance_2020 %>% head())

prices_2019 <- get_average_ticket_price_data()
print(prices_2019 %>% head())

team_names <- attendance_2020 %>% 
  pull(team_name) %>% 
  unique() %>% 
  as.character()
print(team_names %>% head())

team_cities <- sapply(team_names, function(name) {
  attendance_2020 %>% 
    filter(team_name == name) %>% 
    pull(city) %>% 
    .[1]
})
print(team_cities %>% head())

rev_data <- get_revenue_data()
print(rev_data %>% head())

rev_team_data <- get_team_revenue_data(team_names, team_cities)
print(rev_team_data %>% head())
```

### Run models (may take several hours, skip to load models if already done)
```{r run_models, eval=FALSE}
fit_samps <- fit_bayesian_mods() %>% 
  save_fit_tight_priors() %>% 
  save_samps_tight_priors() %>% 
  downsample_samps()
```

### Load models (much faster if models have already been run and samples saved, can be skipped if running model in above chunk)
```{r load_models}
fit_samps <- load_bayesian_samps() %>% downsample_samps()
```

### Load other model metadata
```{r model_info}
prior <- get_priors()[[1]]
print(prior)
var_names <- fit_samps %>% names() %>% .[1:6]
print(var_names)
```

Model: 
$$
\begin{aligned}
\text{Est 2020 Attendance}_\text{team}&\sim\mathcal{N}\left(\mu_\text{team}, \sigma^2\right) \\
\mu_\text{team}&=\alpha_\text{team} \\
      &~~~~~~+\beta_\text{team} \times \text{2019 Year Attendance} \\
      &~~~~~~+\theta \times \text{2019 Num Games Won} \\
\sigma&\sim\text{Gamma}(\alpha_\sigma, \beta_\sigma) \\
\alpha_\text{team}&\sim\mathcal{N}(\alpha, \lambda^2) \\
\alpha&\sim\mathcal{N}(\mu_{\alpha}, \sigma^2_\alpha) \\
\beta_\text{team}&\sim\mathcal{N}(\beta, \eta^2) \\
\beta&\sim\mathcal{N}(\mu_{\beta}, \sigma^2_\beta) \\
\theta&\sim\mathcal{N}(\mu_{\theta}, \sigma^2_\theta) \\
\end{aligned}
$$

## Exploratory Data Analysis

### Who has the most attendees at their home games?
```{r avg_att_eda}
plot_avg_attendance_all_years(attendance, "All teams")
```

Commanders, Cowboys, and Giants have the highest attendance.
Raiders, Cardinals, and Chargers have the lowest.


### Who wins the most?
```{r win_rate_eda}
plot_avg_win_rate_all_years(attendance, "All teams")
```

Patriots, Steelers, and Packers win the most.
Browns, Lions, and Raiders lose the most.

# How much money does the NFL make?
```{r revenue_eda}
plot_yearly_revenue(rev_data, rev_team_data, "All teams")
```

There is a roughly linear trend, increasing by about $1B every year. 
In 2019, NFL revenue totaled $15.26B.

## Parameter modeling

### What is the prior and posterior predicted distribution of the parameters?
```{r parameter_distributions}
for (var in c("overall_alpha", "overall_beta", "theta", "sigma", "alphas", "betas")){
  plt <- plot_distribution(fit_samps, prior, var, team_names, "All teams")
  print(plt)
}
```

### Do the MCMC diagnoistics look healthy (was this a good fit)?
```{r parameter_MCMCdiagnostics}
for (var in c("overall_alpha", "overall_beta", "theta", "sigma", "alphas", "betas")) {
  neff <- get_effective_sample_size(fit_samps, var, team_names, team_name="All teams")
  rhat <- get_rhat(fit_samps, var, team_names, team_name="All teams")
  trace_plt <- get_trace_plot(fit_samps, var, team_names, team_name="All teams")
  acf_plt <- get_acf_plot(fit_samps, var, team_names, team_name="All teams")
  
  print(var)
  print("Effective sample size:")
  print(neff)
  print("Gelman statistic:")
  print(rhat)
  print(trace_plt)
  print(acf_plt)
  print("")
}
```

Fit looks great.
Effective sample size is roughly 10,000 for all parameters.
The Gelman statistic is roughly 1.00 for all parameters.
Nothing concerning is in the trace plots nor the autocorrelation plots.


## Estimated Revenue Loss

### Attendance
Which teams had the biggest attendance loss because of covid?
```{r attendance_results}
post_predicted_attendance <- get_post_predicted_attendance(
  attendance, 
  fit_samps, 
  team_names, 
  "All teams"
)

plot_attendance(
  post_predicted_attendance, 
  attendance_2020, 
  team_names, 
  "All teams"
)
```

The Jets seem to have the worst impact on attendance. 
The Chargers had the smallest impact, because they had such low attendance in the first place.

### Revenue

$$\text{Impact of Covid on Weekly Attendance}_\text{team}\approx\text{Actual 2020 Attendance}_\text{team}-\text{Est 2020 Attendance}_\text{team}$$
$$\text{Est Rev Impact}\approx\sum_{\text{team}\in\text{All teams}}{\text{Impact Weekly Attendance}_\text{team} \times \text{Avg. 2019 Ticket Cost}_\text{team} \times \text{Num Weeks}}$$

How much money was actually lost on ticket sales because of covid? 
```{r revenue_results}
post_predicted_loss_attendance <- get_post_predicted_lost_attendance(
  post_predicted_attendance,
  attendance_2020, 
  team_names, 
  "All teams"
)
print(post_predicted_loss_attendance %>% head())

est_rev_loss <- get_est_lost_rev(
  post_predicted_loss_attendance, 
  prices_2019, 
  "All teams"
)
```

Estimated Revenue Loss: `r est_rev_loss`